On-the-fly Algorithms for GKAT Equivalences \thanks{Identify applicable funding agency here. If none, delete this.}

### 2025-01-16 main.tex
qcfu email
finish proof

\begin{definition}[normalizing bisimulation]\label{def:norm-bisim}
    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a normalizing bisimulation \({≈} ⊆ S × U\) is a GKAT coalgebra that contains \(∼\) with the following transition function:
    \[
        δ_≈((s, u), α) ≜ \begin{cases}
            \accept & \begin{aligned}
                & \text{if } δ_S(s, α) = \accept \\[-3px]
                & \text{and } δ_U(s, α) = \accept
            \end{aligned}\\[5px]
            ((s', u'), p) & 
                \begin{aligned}
                    & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
                    & \text{and } u \transvia{α ∣ p}_{U} u' 
                \end{aligned}\\[5px]
            \reject & \text{otherwise}
        \end{cases}
    \]
    We call \(∼\) \emph{the underlying bisimulation} of \(≈\).
\end{definition}

\begin{lemma}
    Given a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), the following relation \({∼_d} ⊆ S × U\):
    \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}
\end{lemma}

\begin{proof}
    We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:norm-bisim}.
    Specifically, if \(s ∼_d u\) and \(s \transvia{α ∣ p}_{S} s'\) and \(u \transvia{α ∣ p} u'\), then \(s' ∼_d u'\).
    By case analysis on the liveness of \(s'\) and \(u'\) in \(S\) and \(U\):
    \begin{itemize}
        \item If both \(s'\) and \(u'\) are dead, then \(s ∼_d u\) by definition of \(∼_d\).
        \item If both \(s'\) and \(u'\) are live, then \(s, s'\) and \(u, u'\) are live in \(S\) and \(U\) respectively. 
        Thus, \(s \transvia{α ∣ p}_{\norm(S)} s'\), and \(u \transvia{α ∣ p}_{\norm(U)} u'\).
        Therefore, \(s' ∼ u'\) by definition of bisimulation on \(\norm(S)\) and \(\norm(U)\); and \(s ∼_d u'\).
        \item If \(s'\) is live but \(u'\) is dead, then \(s\) is live, and \(s ∼ u\).
        However, \(s \transvia{α ∣ p}_{\norm(S)} s'\) yet \(u \transRej{\norm(U)}{u}\), which violates the condition for bisimulation, thus this case cannot happen.
        \item If \(u'\) is live but \(s'\) is dead, this case also cannot happen similar to the previous case.
    \end{itemize}
\end{proof}

\begin{lemma}\label{thm:norm-bisim-exists-unique}
    For any \(∼\) a normalizing bisimulation with \(∼\) as underlying bisimulation exists and is unique.
\end{lemma}

\begin{proof}
    TODO: sub-coalgebra form a complete lattice.
\end{proof}

@@

\begin{lemma}\label{thm:live-in-component-then-live-in-norm-bisim}
    Given a normalizing bisimulation \(≈\), if a state \((s, u)\) is live in \(≈\), then \(s\) and \(u\) are live in \(S\) and \(U\) respectively, which also implies that \((s, u)\) is live in \(∼\).
\end{lemma}

\begin{proof}
    TODO: by preservation of accepting state and liveness chain.
\end{proof}

\begin{lemma}
    Every state in \((s, u) ∈ {≈} ⊆ S × U\) but not in \({∼} ⊆ \norm(S) × \norm(U)\) then \((s, u)\) is dead in \(≈\).
\end{lemma}

\begin{proof}
    By~\Cref{thm:norm-bisim-not-bisim-then-component-dead}, \(s\) is dead in \(S\), then by contrapositive of~\Cref{thm:live-in-component-then-live-in-norm-bisim}, we obtain that \((s, u)\) is dead in \(≈\).
\end{proof}

\begin{theorem}[Correctness]
    Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a normalizing bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
\end{theorem}

@@

    \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
    Because~\Cref{thm:norm-bisim-exists-unique}, therefore there exists a normalizing bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
    Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
\end{proof}

\begin{theorem}[Recursive Construction]
    Given a relation \({≈} ⊆ S × U\), \(≈\) is a carrier of some normalizing bisimulation if and only if all the following condition are satisfied: for all pair of states \(s ≈ u\),
    \begin{enumerate}
        \item\label{itm:acc-condition} for all \(α ∈ \At\), 
        \(s \transAcc{S}{α}\) if and only if \(u \transAcc{U}{α}\);
        \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≈ u'\).
        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(u'\) are dead.
    \end{enumerate}
\end{theorem}

\begin{proof}
    To show the \(⟹\) direction, assuming there exists a normalizing bisimulation \(≈\), then it would satisfy all the conditions. 
    We first note that the pairs \((s, u) ∈ {≈}\) can satisfy one of two properties (or both), either \(s\) and \(u\) are both dead in \(S\) and \(U\), or \((s, u)\) is in the underlying bisimulation \(∼\).
    Then we will verify that all the condition is satisfied both when \(s ∼ u\) and when \(s, u\) are dead.

@@

    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of normalizing bisimulation (\Cref{def:norm-bisim}).

@@

    Then we prove the \(⟸\) direction.
    TODO: by constructing a \(≈\) and show that \(\norm(≈)\) is a bisimulation. Maybe using the maximal bisimulation?
\end{proof}

\begin{lemma}\label{thm:bisim-between-dead}
    Given any two dead states \(s ∈ S\) and \(u ∈ U\), and the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), then we have \(s ≡ u\).
\end{lemma}

\begin{theorem}[Equivalence Relation]
    Given two states in GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), there exists a normalizing bisimulation \(≈\) between \(S\) and \(U\) s.t. \(s ≈ u\), if and only if there exists a normalizing bisimulation that is an \emph{equivalence relation} \(≊\) in \(S + U\) s.t. \(s ≊ u\).
\end{theorem}

\begin{proof}
    First, recall that by~\Cref{thm:} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
    Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra~\Cref{thm:norm-sub-coalg}, we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).

    The \(⟹\) direction. 
    Because sub-coalgebra preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}), then \(\norm(≈)\) is a bisimulation between \(\norm(S + U)\) and itself.
    We consider \(≡\) to be the maximal bisimulation between \(\norm(S + U)\) and itself, then \(≡\) is an equivalence relation~\cite[]{}.

    Let \(≊\) be the normalizing bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
    All the \((s, u) ∈ {≊}\) are either \(s ≡ u\) or \(s\) and \(u\) are dead in \(S\) and \(U\), which implies that both are dead in \(S + U\).
    However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
    Finally, by definition of normalizing bisimulation (\Cref{def:norm-bisim}), \({≡} ⊆ {≊}\).
    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a normalizing bisimulation that is also an equivalence relation, which also contains \((s, u)\):
    \[(s, u) ∈ {≈} = \norm(≈) ⊆ {≡} ⊆ {≊},\]
    where all the \(=\) and \(⊆\) are orders on the carrier set, instead of the coalgebra.

    The \(⟸\) direction can be split into two cases, by whether \((s, u)\) is in the underlying bisimulation of \(≊\), which we denote as \(≃\): 
    \begin{itemize}
        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
        Taking the normalizing bisimulation over \(∼\) will yield a desired normalizing bisimulation \(≈\) s.t. \(s ≈ u\).
        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:norm-bisim-not-bisim-then-component-dead}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
        Then we take \(≈\) to be normalizing bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
        \qedhere
    \end{itemize}
\end{proof}

@@

% % \begin{definition}[normalizing bisimulation]\label{def:norm-bism}
% %     A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a normalizing bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
% %     \begin{enumerate}
% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
% %     \end{enumerate}
% %     And \(≃\) can be endowed with the following GKAT coalgebra structure:
% %     \[
% %         δ_≃((s, u), α) ≜ \begin{cases}
% %             \accept & \begin{aligned}
% %                 & \text{if } δ_S(s, α) = \accept \\[-3px]
% %                 & \text{and } δ_U(s, α) = \accept
% %             \end{aligned}\\[5px]
% %             ((s', u'), p) & 
% %                 \begin{aligned}
% %                     & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
% %                     & \text{and } u \transvia{α ∣ p}_{U} u' 
% %                 \end{aligned}\\[5px]
% %             \reject & \text{otherwise}
% %         \end{cases}
% %     \]
% % \end{definition}

% \begin{lemma}[Perservation of Liveness]\label{thm:norm-bisim-perserve-liveness}
%     For any normalizing bisimulation \(≃\) and a pair of states \(s ≃ u\), \((s, u)\) is live in \(≃\) if and only if \(s\) is live in \(S\), and similarly for \(u\) in \(U\). 
%     This implies that if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
% \end{lemma}

% \begin{proof}
%     First by definition of normalizing bisimulation that a state \(s\) is accepting in \(S\) if and only if \((s, u)\) is accepting in \(≃\).

@@

% \begin{theorem}
%     Given a normalizing bisimulation \(≃: S × U\), its normalization is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\)
% \end{theorem}

@@

>>>>

\begin{definition}[greedy bisimulation]\label{def:norm-bisim}
    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \({≈} ⊆ S × U\) is a GKAT coalgebra that contains \(∼\) with the following transition function:
    \[
        δ_≈((s, u), α) ≜ \begin{cases}
            \accept & \begin{aligned}
                & \text{if } δ_S(s, α) = \accept \\[-3px]
                & \text{and } δ_U(s, α) = \accept
            \end{aligned}\\[5px]
            ((s', u'), p) & 
                \begin{aligned}
                    & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
                    & \text{and } u \transvia{α ∣ p}_{U} u' 
                \end{aligned}\\[5px]
            \reject & \text{otherwise}
        \end{cases}
    \]
    We call \(∼\) \emph{the underlying bisimulation} of \(≈\).
\end{definition}

\begin{lemma}\label{thm:dead-construction-greedy-bisim}
    Given any eager bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
    \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}, which implies \({≈} ⊑ \)
    Furthermore, \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
\end{lemma}

\begin{proof}
    We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:norm-bisim}.
    Specifically, if \(s ∼_d u\) and \(s \transvia{α ∣ p}_{S} s'\) and \(u \transvia{α ∣ p} u'\), then \(s' ∼_d u'\).
    By case analysis on the liveness of \(s'\) and \(u'\) in \(S\) and \(U\):
    \begin{itemize}
        \item If both \(s'\) and \(u'\) are dead, then \(s ∼_d u\) by definition of \(∼_d\).
        \item If both \(s'\) and \(u'\) are live, then \(s, s'\) and \(u, u'\) are live in \(S\) and \(U\) respectively. 
        Thus, \(s \transvia{α ∣ p}_{\norm(S)} s'\), and \(u \transvia{α ∣ p}_{\norm(U)} u'\).
        Therefore, \(s' ∼ u'\) by definition of bisimulation on \(\norm(S)\) and \(\norm(U)\); and \(s ∼_d u'\).
        \item If \(s'\) is live but \(u'\) is dead, then \(s\) is live, and \(s ∼ u\).
        However, \(s \transvia{α ∣ p}_{\norm(S)} s'\) yet \(u \transRej{\norm(U)}{u}\), which violates the condition for bisimulation, thus this case cannot happen.
        \item If \(u'\) is live but \(s'\) is dead, this case also cannot happen similar to the previous case.
    \end{itemize}

    By~\Cref{thm:bisim-between-dead}, we can construct a bisimulation \(∼_{s,u} ≜ \{(s, u)\} ⊆ \norm(S) × \norm(U)\) between any two dead states \(s ∈ S\) and \(u ∈ U\).
    Then because bisimulation is closed under union~\cite[Theorem 5.5]{rutten_UniversalCoalgebraTheory_2000}, \(\norm(∼_d) = {∼} ∪ ⋃ \{∼_{s, u} ∣ \text{\(s, u\) are dead}\}\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
    TODO: bisimulation are uniquely determined by its carrier.
\end{proof}

\begin{lemma}\label{thm:norm-bisim-exists-unique}
    For any bisimulation between two normalized GKAT coalgebra \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
\end{lemma}

\begin{proof}
    Because sub-coalgebra form a complete lattice under union and intersection~\cite[Theorem 6.4]{rutten_UniversalCoalgebraTheory_2000}, then we can construct \(≈\) by taking the intersection of all the sub-coalgebra of \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}.
\end{proof}

@@

\begin{theorem}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
    Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then the following are equivalent
    \begin{itemize}
        \item \((s, u)\) is live (or dead) in \({≈}\);
        \item \(s\) is live (or dead) in \(S\);  
        \item \(u\) is live (or dead) in \(U\).
    \end{itemize}
    Furthermore, if \((s, u)\) is in the underlying bisimulation \(∼\) of \(≈\), then all the above is also equivalent to \((s, u)\) is live (or dead) in \(∼\).
\end{theorem}

\begin{proof}
    Let \(∼\) by the underlying bisimulation of \(≈\), consider the \(∼_d\) in~\Cref{thm:dead-construction-greedy-bisim}, then we know that \({≈} ⊑ {∼_d}\).

    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then
    \begin{align*}
        & \text{\(s\) is live in \(S\)} \\
        \text{ iff } & \text{\(s\) is live in \(\norm(S)\)} 
            & \text{\Cref{thm:norm-perserve-liveness}}\\  
        \text{ iff } & \text{\((s, u)\) is live in \(\norm(∼_d)\)} 
            & \text{\Cref{thm:bisim-preserve-liveness}}\\
        \text{ iff } & \text{\((s, u)\) is live in \(∼_d\)} 
            & \text{\Cref{thm:norm-perserve-liveness}}\\
        \text{ iff } & \text{\((s, u)\) is live in \(≈\)} 
            & \text{\Cref{thm:sub-coalg-preserve-liveness}}
    \end{align*}
\end{proof}

\begin{lemma}\label{thm:greedy-bisim-iff-norm-bisim}
    Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:norm-bisim}.
    \(≈\) is a greedy bisimulation if and only if \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\)
\end{lemma}

\begin{proof}
    The \(⟹\) direction, we let \(∼_d\) be defined as in~\Cref{thm:dead-construction-greedy-bisim}.
    Because \({≈} ⊑ {∼_d}\) and~\Cref{thm:norm-sub-coalg}, \(\norm(≈) ⊑ \norm(∼_d)\). 
    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then we can show that \(\norm(≈)\) is a bisimulation by factoring through \(\norm(∼_d)\), in the category of GKAT coalgebras:
    \begin{align*}
        & \norm(≈) 
        \xrightarrow{i} \norm(∼_d) 
        \xrightarrow{π₁} \norm(S) \\
        & \norm(≈) 
        \xrightarrow{i} \norm(∼_d) 
        \xrightarrow{π₂} \norm(U) 
    \end{align*}
    because inclusion homomorphism \(i\) do not change its input, the two homomorphism obtained by bisimulation is indeed \(π₁: \norm(≈) → \norm(S)\) and \(π₂: \norm(≈) → \norm(U)\), implying both projections of \(\norm(≈)\) are homomorphism, meaning \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).

    The \(⟸\) direction can be proven by simply letting \(\norm(≈)\) be the underlying bisimulation of \(≈\), because \(\norm(≈)\) has the same carrier set of \(≈\), then \(≈\) has to be the least GKAT coalgebra with \(δ_≈\) that contains \(\norm(≈)\), satisfying all the conditions for greedy bisimulation.
\end{proof}

\begin{theorem}[Correctness]
    Given two states in GKAT coalgebras \(s ∈ S\) and \(u ∈ ,U\), there exists a greedy bisimulation \({≈} ⊆ S × U\) s.t. \(s ≈ u\), if and only if \(⟦s⟧_S = ⟦u⟧_U\).
\end{theorem}

@@

    \(⟸\) direction, because \(⟦s⟧_S = ⟦u⟧_U\), then by~\cref{thm:norm-bisim-correctness}, then there a bisimulation \(∼ ⊆ \norm(S) × \norm(U)\) s.t. \(s ∼ u\). 
    Because~\Cref{thm:norm-bisim-exists-unique}, therefore there exists a greedy bisimulation \(≈\) with \(∼\) as its underlying bisimulation. 
    Because the carriers \({∼} ⊆ {≈}\), therefore \(s ∼ u\) implies \(s ≈ u\).
\end{proof}

\begin{theorem}[Recursive Construction]\label{thm:recursive-construction}
    Given a relation \({≈} ⊆ S × U\), \(≈\) is a carrier of some greedy bisimulation if and only if all the following condition are satisfied: for all pair of states \(s ≈ u\),
    \begin{enumerate}
        \item\label{itm:acc-condition} for all \(α ∈ \At\), 
        \(s \transAcc{S}{α}\) if and only if \(u \transAcc{U}{α}\);
        \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
        \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≈ u'\).
        \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(u'\) are dead.
    \end{enumerate}
\end{theorem}

\begin{proof}
    To show the \(⟹\) direction, assuming there exists a greedy bisimulation \(≈\), then it would satisfy all the conditions. 
    We first note that the pairs \((s, u) ∈ {≈}\) can satisfy one of two properties (or both), either \(s\) and \(u\) are both dead in \(S\) and \(U\), or \((s, u)\) is in the underlying bisimulation \(∼\).
    Then we will verify that all the condition is satisfied both when \(s ∼ u\) and when \(s, u\) are dead.

@@

    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of greedy bisimulation (\Cref{def:norm-bisim}).

@@

    Then we prove the \(⟸\) direction.
    If \(≈\) satisfy all the conditions, then by condition~\labelcref{itm:transition-bisim}, \(≈\) is closed under \(δ_≈\), hence is a GKAT coalgebra.
    We let its underlying bisimulation be \(\norm(≈)\), which is sound if \(\norm(≈)\) is a bisimulation (\cref{thm:greedy-bisim-iff-norm-bisim}).
    Thus, we only need to show that \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\). Without loss of generality, we will show that \(π₁: \norm(≈) → \norm(S)\) is a bisimulation.

    For all pair of states \((s, u) ∈ \norm(≈)\) and any atom \(α ∈ \At\), we will verify all the conditions of a homomorphism.
    For the accepting case, by the definition of \(δ_{≈}\):
    \begin{align*}
        (s, u) \transAcc{\norm(≈)}{α}
        & \text{ iff } (s, u) \transAcc{≈}{α} \\  
        & \text{ implies } s \transAcc{≈}{α}
    \end{align*}
    For the transition case, because \((s, u)\) is live in \(≈\) implies that \(s\) is live in \(S\) (\Cref{thm:greedy-bisim-perserve-liveness}):
    \begin{align*}
        & (s, u) \transvia{α ∣ p}_{\norm(S)} (s', u') \\
        & \text{iff }
            (s, u) \transvia{α ∣ p}_{S} (s', u')
            \text{ and \(s, u\) is live in \(≈\)} \\  
        & \text{implies }
            s \transvia{α ∣ p}_S s' 
            \text{ and \(s\) is live in \(S\)} \\  
        & \text{implies }
            s \transvia{α ∣ p}_{\norm(S)} s' 
    \end{align*}
    Finally, the rejection case: \((s, u) \transRej{\norm(≈)}{α}\) either when \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) and \((s', u')\) is dead, or \((s, u) \transRej{≈}{α}\).
    We will show that both cases will lead to \(s \transRej{\norm(S)}{α}\).
    The \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) case can be proven by the definition of \(δ_≈\) and~\Cref{thm:greedy-bisim-perserve-liveness}:
    \begin{align*}
        & (s, u) \transvia{α ∣ p}_{≈} (s', u') 
            \text{ and \((s', u')\) is dead in \(≈\)} \\
        \text{implies } & s \transvia{α ∣ p}_{S} s' 
            \text{ and \(s'\) is dead in \(S\)} \\  
        \text{implies } & s \transRej{\norm(S)}{α}.
    \end{align*}
    Then the hardest case is when \((s, u) \transRej{≈}{α}\) because by definition of \(δ_≈\) (\Cref{def:norm-bisim}) there are many cases that can lead to rejection.
    We will need to eliminate irrelevant cases using the condition provided by the construction.
    We proceed by case analysis on dynamics of \(s\) and \(u\) on \(α\); first listing some cases that contradicts the premise \((s, u) \transRej{≈}{α}\):
    \begin{itemize}
        \item If \(s\) accepts \(α\) in \(S\), then \(u\) necessarily accepts \(α\) in \(U\) by condition~\labelcref{itm:acc-condition}, which implies \((s, u) \transAcc{≈}{α}\), contradicting the premise.
        \item Similarly, if \(u\) accepts \(α\) in \(U\), then \((s, u) \transAcc{≈}{α}\), contradicting the premise.
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\) then \((s, u) \transvia{α ∣ p}_≈ (s', u;)\), contradicting the premise.
    \end{itemize}
    Then we list the other cases, which indeed satisfy \((s, u) \transRej{≈}{α}\), and show that all of them leads to \(s \transRej{\norm(S)}{α}\), using the condition in the theorem:
    \begin{itemize}
        \item If \(s \transRej{S}{α}\), then \(s \transRej{\norm(S)}{α}\).
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transRej{U}{α}\), then by condition~\labelcref{itm:rej-or-dead}, \(s'\) is dead, and \(s \transRej{\norm(S)}{α}\).
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ q}_U u'\), then by condition~\labelcref{itm:transition-dead}, both \(s'\) and \(u'\) are dead, and \(s \transRej{\norm(S)}{α}\).
        \qedhere
    \end{itemize}
\end{proof}

\begin{lemma}\label{thm:bisim-between-dead}
    Given any two dead states \(s ∈ S\) and \(u ∈ U\), we can construct a bisimulation \(∼\) between \(\norm(S)\) and \(\norm(U)\):
    \begin{align*}
        ∼ &≜ \{(s, u)\}, & δ_∼((s, u)) & ≜ \reject.
    \end{align*}
    Therefore, for the maximal bisimulation \({≡} ⊆ \norm(S) × \norm(U)\), \(s ≡ u\).
\end{lemma}

\begin{theorem}[Equivalence Relation]
    Given two states in GKAT coalgebra \(s ∈ S\) and \(u ∈ U\), there exists a greedy bisimulation \(≈\) between \(S\) and \(U\) s.t. \(s ≈ u\), if and only if there exists a greedy bisimulation that is an \emph{equivalence relation} \(≊\) in \(S + U\) s.t. \(s ≊ u\).
\end{theorem}

\begin{proof}
    First, recall that by~\Cref{thm:greedy-bisim-iff-norm-bisim} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
    Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra~\Cref{thm:norm-sub-coalg}, we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).

    The \(⟹\) direction. 
    Because sub-coalgebra preserves bisimulation (\Cref{thm:sub-coalg-preserve-bisim}), then \(\norm(≈)\) is a bisimulation between \(\norm(S + U)\) and itself.
    We consider \(≡\) to be the maximal bisimulation between \(\norm(S + U)\) and itself, then \(≡\) is an equivalence relation~\cite[Corollary 5.6]{rutten_UniversalCoalgebraTheory_2000}.

    Let \(≊\) be the greedy bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
    All the \((s, u) ∈ {≊}\) are either \(s ≡ u\) or \(s\) and \(u\) are dead in \(S\) and \(U\), which implies that both are dead in \(S + U\).
    However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
    Finally, by definition of greedy bisimulation (\Cref{def:norm-bisim}), \({≡} ⊆ {≊}\).
    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a greedy bisimulation that is also an equivalence relation, which also contains \((s, u)\):
    \[(s, u) ∈ {≈} = \norm(≈) ⊆ {≡} ⊆ {≊},\]
    where all the \(=\) and \(⊆\) are orders on the carrier set, instead of the coalgebra.

    The \(⟸\) direction can be split into two cases, by whether \((s, u)\) is in the underlying bisimulation of \(≊\), which we denote as \(≃\): 
    \begin{itemize}
        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
        Taking the greedy bisimulation over \(∼\) will yield a desired greedy bisimulation \(≈\) s.t. \(s ≈ u\).
        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:norm-bisim-not-bisim-then-component-dead}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
        Then we take \(≈\) to be greedy bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
        \qedhere
    \end{itemize}
\end{proof}

@@

% % \begin{definition}[greedy bisimulation]\label{def:norm-bism}
% %     A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a greedy bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
% %     \begin{enumerate}
% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
% %     \end{enumerate}
% %     And \(≃\) can be endowed with the following GKAT coalgebra structure:
% %     \[
% %         δ_≃((s, u), α) ≜ \begin{cases}
% %             \accept & \begin{aligned}
% %                 & \text{if } δ_S(s, α) = \accept \\[-3px]
% %                 & \text{and } δ_U(s, α) = \accept
% %             \end{aligned}\\[5px]
% %             ((s', u'), p) & 
% %                 \begin{aligned}
% %                     & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
% %                     & \text{and } u \transvia{α ∣ p}_{U} u' 
% %                 \end{aligned}\\[5px]
% %             \reject & \text{otherwise}
% %         \end{cases}
% %     \]
% % \end{definition}

% \begin{lemma}[Perservation of Liveness]\label{thm:norm-bisim-perserve-liveness}
%     For any greedy bisimulation \(≃\) and a pair of states \(s ≃ u\), \((s, u)\) is live in \(≃\) if and only if \(s\) is live in \(S\), and similarly for \(u\) in \(U\). 
%     This implies that if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
% \end{lemma}

% \begin{proof}
%     First by definition of greedy bisimulation that a state \(s\) is accepting in \(S\) if and only if \((s, u)\) is accepting in \(≃\).

@@

% \begin{theorem}
%     Given a greedy bisimulation \(≃: S × U\), its normalization is indeed a bisimulation between \(\norm(S)\) and \(\norm(U)\)
% \end{theorem}

@@

\begin{theorem}\label{thm:greedy-bisim-one-dead}

\end{theorem}
